---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```
```{r}
library(dplyr)
library(tidyverse)
library(stringr)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
deaths <- av |>
  pivot_longer(cols = starts_with("Death"),
               names_to = "Time",
               values_to = "Death") |>
  mutate(Time = parse_number(str_remove(Time,"Death")))
deaths
```

Similarly, deal with the returns of characters.

```{r}
returns <- av |>
  pivot_longer(cols = starts_with("Return"),
               names_to = "Time",
               values_to = "Return") |>
  mutate(TimeReturn = parse_number(str_remove(Time,"Return")))
returns
```

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
#Average number of deaths
sum(deaths$Death == "YES") / length(unique(av$Name.Alias))
```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.


## Michael Friedman
### FiveThirtyEight Statement

> "But you can only tempt death so many times. There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife".

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

```{r}

# Find how many Avengers died and remove duplicate observations
dead_avengers <- deaths %>%
  filter(Death == "YES") %>%
  distinct(URL, .keep_all = TRUE) %>%
  nrow()

# Find how many Avengers returned from their first death
returned_avengers <- deaths %>%
  filter(Death == "YES") %>%
  distinct(URL, .keep_all = TRUE) %>%
  filter(Return1 == "YES") %>%
  nrow()

# Calculate and print the proportion
did_they_return <- returned_avengers / dead_avengers
print(did_they_return)

```

### Include your answer
Michael: The propotion of Avengers who came back to life after their first death was in fact 2/3.

## Subham Bhattacharya
### FiveThirtyEight Statement
> Subham Statment: "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team. That’s about 40 percent of all people who have ever signed on to the team."

### Subham code
```{r}
# Calculate the total number of Avengers who died at least once
avengers_died <- deaths %>%
  filter(Death == "YES") %>%
  distinct(Name.Alias) %>%
  nrow()

# Calculate the percentage of Avengers who died at least once
total_avengers <- length(unique(av$Name.Alias))
percentage_died <- (avengers_died / total_avengers) * 100

cat("Out of", total_avengers, "listed Avengers, my analysis found that", avengers_died, "had died at least one time after they joined the team. That’s about", round(percentage_died, 2), "percent of all people who have ever signed on to the team.\n")

```

### Subham answer
Subham: Based on my code and analysis I found that my answer is slightly different from the provided statement in my code I got that there are a total of 163 avengers among which 63 have died atleast once, which is about 39.26% so my analysis gets slightly different result from the provided the result.



## Manjul Balayar

### FiveThirtyEight Statement

> Let's check if the Hulk has died less often than Thor

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

```{r}
specific_avengers <- av %>% 
  filter(grepl("Robert Bruce Banner|Thor Odinson", Name.Alias))

deaths_specific <- specific_avengers |>
  pivot_longer(cols = starts_with("Death"),
               names_to = "Time",
               values_to = "Death") |>
  mutate(Time = parse_number(Time)) %>%
  group_by(Name.Alias) %>%
  summarise(Total_Deaths = sum(Death == "YES", na.rm = TRUE))

print(deaths_specific)

```

### Include your answer

Manjul: After analyzing the data, it was found that Hulk has faced death once, where as Thor has faced death twice. Meaning Hulk has died less than Thor.

Upload your changes to the repository. Discuss and refine answers as a team.
